---
title: "Homework 2"
author: "Jessica Lavery"
date: "9/27/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

# Problem 1

Mr. Trash Wheel dataset
```{r}
# read in Mr. Trash Wheel dataset
trash_wheel <- read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = "Mr. Trash Wheel",
                          range = "A2:N336") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls_int = as.integer(round(sports_balls, 1)))

# check the above
# class(pull(trash_wheel, sports_balls_int))
```

Precipitation data, 2017-2018
```{r}
# read in 2017 precipitation data
precip_2017 <- read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = "2017 Precipitation",
                          range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  rename(precip = total) %>% 
  mutate(year = 2017)

# read in 2018 precipitation data
precip_2018 <- read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = "2018 Precipitation",
                          range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  rename(precip = total) %>% 
  mutate(year = 2018)

# combine the precipitation datasets 
precip_2017_2018 <- bind_rows(precip_2017, precip_2018) %>% 
  mutate(month = month.name[month])
```

Precipitation data was recorded for `r nrow(precip_2017_2018 %>% filter(year == 2017))` months in 2017 and `r nrow(precip_2017_2018 %>% filter(year == 2018))` months in 2018. The average precipitaiton among months recorded across both years was `r mean(pull(precip_2017_2018, precip))` inches (standard deviation `r round(sd(pull(precip_2017_2018, precip)),2)`), ranging from `r min(pull(precip_2017_2018, precip))` to `r max(pull(precip_2017_2018, precip))`. The total precipitation was `r sum(pull(precip_2017_2018, precip))` in 2017 and `r mean(pull(precip_2017_2018, precip))` in 2018. 