---
title: "Homework 2"
author: "Jessica Lavery"
date: "9/27/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(kableExtra)

```

# Problem 1

Mr. Trash Wheel dataset
```{r}
# read in Mr. Trash Wheel dataset
trash_wheel <- read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = "Mr. Trash Wheel",
                          range = "A2:N336") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls_int = as.integer(round(sports_balls, 1)))

# check the above
# class(pull(trash_wheel, sports_balls_int))
```

Precipitation data, 2017-2018
```{r}
# read in 2017 precipitation data
precip_2017 <- read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = "2017 Precipitation",
                          range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  rename(precip = total) %>% 
  mutate(year = 2017)

# read in 2018 precipitation data
precip_2018 <- read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                          sheet = "2018 Precipitation",
                          range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  rename(precip = total) %>% 
  mutate(year = 2018)

# combine the precipitation datasets 
precip_2017_2018 <- bind_rows(precip_2017, precip_2018) %>% 
  mutate(month = month.name[month])
```

The trash wheel dataset records trash found in the Harbor by month and year, from `r min(pull(trash_wheel, date))` through `r max(pull(trash_wheel, date))`. The amount of trash is recorded in terms of weight, volume, as well as by type of trash: plastic bottles, polystyrene, cigarette butts, glass bottles, grocery bags, chip bags, and sports balls. The total amount of trash over the study period was `r sum(pull(trash_wheel, weight_tons))` tons, or `r sum(pull(trash_wheel, volume_cubic_yards))` cubic yards. The median number of sports balls in a dumpster in 2017 was `r median(pull(trash_wheel %>% filter(year == "2017")))`.

The precipitation datasets record the amount of precipitation by month and year. Precipitation data was recorded for `r nrow(precip_2017)` months in 2017 and `r nrow(precip_2018)` months in 2018. The average precipitaiton among months recorded across both years was `r mean(pull(precip_2017_2018, precip))` inches (standard deviation `r round(sd(pull(precip_2017_2018, precip)),2)`), ranging from `r min(pull(precip_2017_2018, precip))` to `r max(pull(precip_2017_2018, precip))`. The total precipitation was `r sum(pull(precip_2017, precip))` in 2017 and `r mean(pull(precip_2018, precip))` in 2018. 

# Problem 2
```{r}
# read in and clean pols-month data
pols <- read_csv("./data/pols-month.csv") %>% 
  separate(col = mon, into = c("year", "month", "day"), sep = '-') %>% 
  mutate(month = month.abb[as.numeric(month)],
         year = as.numeric(year), 
         president = case_when(
           prez_gop == 1 ~ "gop",
           prez_dem == 1 ~ "dem"
         )) %>% 
  select(-prez_gop, -prez_dem, -day)

pols

# read in and clean snp data
snp <- read_csv("./data/snp.csv") %>% 
  separate(col = date, into = c("month", "day", "year"), sep = '/') %>% 
  arrange(year, as.numeric(month)) %>% 
  mutate(month = month.abb[as.numeric(month)],
         year = as.numeric(year)) %>% 
  select(year, month, close)

snp

# read in and tidy unemployment data
unemp <- read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(cols = jan:dec,
               names_to = "month",
               values_to = "unemp_rate") %>% 
  mutate(month = str_to_sentence(month))

unemp

# merge snp and pols
pols_snp <- inner_join(pols, snp, by = c("year", "month"))

pols_snp_unemp <- inner_join(pols_snp, unemp, by = c("year", "month"))
```

The pols dataset contains information about the number of national politicians in each of the two primary parties (democratic, republican) by month and year (`r min(pull(pols, year))` - `r max(pull(pols, year))`). The second data source in problem two (snp) contains the closing Standard and Poor's (S&P) stock market index on the first of the month between `r min(pull(snp, year))` and `r max(pull(snp, year))`. The third dataset of interest (unemp) contains the percentage of unemployment by month from `r min(pull(unemp, year))` through `r max(pull(unemp, year))`.  

Combining these datasets among years and months that all information is provided (political information, S&P index, and unemployment), we have data for `r min(pull(pols_snp_unemp, year))` - `r max(pull(pols_snp_unemp, year))`. Key information contained in the combined dataset is stored in the variables `r names(pols_snp_unemp)[3:9]` (political), `r names(pols_snp_unemp)[10]` (S&P index at market close), and `r names(pols_snp_unemp)[11]` (unemployment). 

Over this time period, there were `r nrow(pols_snp_unemp %>% filter(president=="dem"))` months with a democratic president and `r nrow(pols_snp_unemp %>% filter(president=="gop"))` months with a republican president. Under democratic presidents, the S&P closing index ranged from `r round(min(pull(pols_snp_unemp %>% filter(president=="dem"), close)), 2)` to `r round(max(pull(pols_snp_unemp %>% filter(president=="dem"), close)), 2)` and unemployment ranged from `r min(pull(pols_snp_unemp  %>% filter(president=="dem"), unemp_rate))`% to `r max(pull(pols_snp_unemp %>% filter(president=="dem"), unemp_rate))`%. Whereas under republican presidents, the S&P closing index ranged from `r round(min(pull(pols_snp_unemp %>% filter(president=="gop"), close)), 2)` to `r round(max(pull(pols_snp_unemp %>% filter(president=="gop"), close)), 2)` and unemployment ranged from `r min(pull(pols_snp_unemp %>% filter(president=="gop"), unemp_rate))`% to `r max(pull(pols_snp_unemp %>% filter(president=="gop"), unemp_rate))`%. 

# Problem 3
```{r}
# read in and clean popular baby names data, removing duplicate rows
baby_names <- read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>% 
  mutate(gender = str_to_sentence(gender),
         ethnicity = str_to_title(ethnicity),
         childs_first_name = str_to_sentence(childs_first_name),
         ethnicity = recode(ethnicity, 
                            "White Non Hisp" = "White Non Hispanic",
                            "Black Non Hisp" = "Black Non Hispanic",
                            "Asian And Paci" = "Asian And Pacific Islander")) %>% 
  distinct() %>% 
  arrange(year_of_birth)
```

The below table shows the rank of the name Olivia by year and ethnicity.
```{r}
# table of popularity (rank) of the name Olivia by year
olivia <- baby_names %>% 
  filter(childs_first_name == "Olivia") %>% 
  select(-gender, -childs_first_name, -count) %>% 
  pivot_wider(names_from = year_of_birth,
              values_from = rank)

olivia %>% kable()
```

The below table shows the most popular baby name among male children by year and ethnicity.
```{r}
# most popular name among male children over time
popular_male <- baby_names %>% 
  filter(rank == 1, gender == "Male") %>% 
  select(-gender, -count, -rank) %>% 
  pivot_wider(names_from = year_of_birth,
              values_from = childs_first_name)

popular_male %>% kable()
```

The following scatterplot shows the number of male, white non-hispanic children born in 2016 with a given name against the rank in popularity in that name. 

```{r}
# male, white non-hispanic children born in 2016
male_white_2016 <- baby_names %>% 
  filter(year_of_birth == 2016, gender == "Male", ethnicity == "White Non Hispanic")

ggplot(data = male_white_2016, aes(x = rank, y = count)) +
  geom_point() + 
  xlab("Rank of name") +
  ylab("Number of children with name")
```

